---
title: "Dataframe Exercies"
subtitle: "Data 180, Professor Kessler"
author: 
  name: "My Name"
  email: "xxxx@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains a set of exercises which we will complete both in class and at home. Please attempt each of these exercises at your own pace.

### First steps...

Create a data frame named `students` with the following columns:
- `Name`: "Alice", "Bob", "Charlie", "David", "Eva"
- `Age`: 18, 21, 20, 19, 18
- `Major`: "Math", "Physics", "Biology", "Chemistry", "Math"
- `FirstYear`:2024, 2021, 2022, 2023, 2024

# Exercise 1: Create the data frame
```{r}

```

# Exercise 2: Count rows, columns, and dimensions
```{r}

```
# Exercise 3: Filter for Math majors
```{r}

```
# Exercise 4: Filter for students older than 22
```{r}

```
# Exercise 5: Filter for Math majors older than 22
```{r}

```
# Exercise 6: Add GraduationYear column
```{r}

```
# Exercise 7: Creating a Vector
```{r}

```
# Exercise 8: Indices
```{r}

```
# Exercise 9: Creating functions
```{r}

```
