
### **Practice Exam: DATA 180-01**
---
title: "DATA 180-01 Exam Practice"
subtitle: "Fall 2024, Professor Kessler"
author: 
  name: "Your Name"
  email: "XXXX@dickinson.edu"
date: "`r format(Sys.Date(), "%B %d, %Y")`"
output: 
  html_document
---

## Problem 1
(Each part worth 3 points)

Create a vector named `departments` with the names of different company departments: "Finance", "HR", "IT", "Marketing", "Operations"

```{r}

```

a) The `Marketing` department has been renamed to `Sales and Marketing`. Update the vector to reflect this change.

```{r}
# Replace the incorrect entry in the vector

```

b) Write a function named `print_departments` that takes a vector of departments and prints each department along with its position in the vector (e.g., "1: Finance").

```{r}
# Create the function to print each department and its position

```

c) Create a data frame named `department_budget` that contains the following information:

|Department         | Budget ($) in Millions |
|:------------------|:-----------------------|
|Finance            | 5.0                    |
|HR                 | 2.0                    |
|IT                 | 8.0                    |
|Sales and Marketing| 4.5                    |
|Operations         | 6.0                    |

```{r}
# Create the data frame

```

d) The budget of the `IT` department has increased by 1.5 million. Update the `department_budget` data frame accordingly.

```{r}
# Update the budget of IT department

```

e) Calculate the total budget of all departments.

```{r}
# Calculate the total budget

```
f) Calculate the mean and median budget.

```{r}
# Calculate the total budget

```


---

## Problem 2
(Each part worth 4 points)

Consider the sales data of a company for two products, `Product_A` and `Product_B`, for each quarter of the year:

```{r}
sales_data <- data.frame(
  Quarter = c("Q1", "Q2", "Q3", "Q4"),
  Product_A = c(15000, 22000, 18000, 24000),
  Product_B = c(18000, 20000, 21000, 23000)
)
sales_data
```

a) Create a new column named `Total_Sales` that contains the total sales for both products in each quarter.

```{r}
# Create the Total_Sales column

```

b) Create a new column named `Performance` with values "Good" if total sales exceed 40000, otherwise "Needs Improvement".

```{r}
# Create the Performance column
```

c) Plot a bar chart that shows the total sales of `Product_A` and `Product_B` across all quarters combined.

```{r}
# Plot the bar chart
```

d) Write a function named `quarter_summary` that takes the sales data and returns the average sales of `Product_A` and `Product_B`.

```{r}
# Create the function to calculate average sales
```

---

## Problem 3
(Each part worth 5 points)

We have two data frames: `students` with student names and their majors, and `courses` with course details:

```{r}
students <- data.frame(
  Student = c('Alice', 'Bob', 'Charlie', 'David', 'Eve'),
  Major = c('Mathematics', 'Physics', 'Economics', 'Computer Science', 'Biology')
)

courses <- data.frame(
  Course = c('Calculus', 'Physics 101', 'Econometrics', 'Programming 101', 'Genetics'),
  Major = c('Mathematics', 'Physics', 'Economics', 'Computer Science', 'Biology')
)
```



a) Create a variable named `course_difficulty` for each course with levels: "Easy", "Medium", "Hard". Assign them as follows: `Calculus` - "Medium", `Physics 101` - "Easy", `Econometrics` - "Hard", `Programming 101` - "Easy", `Genetics` - "Medium".

```{r}
# Create the factor variable

```

b) Write a function named `course_summary` that takes a data frame of courses and returns the number of courses for each difficulty level.

```{r}
# Create the function to count courses by difficulty level

```


---

### Practice Questions

Can a variable be both continuous and discrete?

If a variable `weight_kg` contains a numeric value representing weight with up to two decimal places, what type of variable is `weight_kg`?

What is the main purpose of the `==` operator in R?

Write a single line of R code that returns the following vector: (1, 4, 7, 10, 13). Do not simply write out `c(1, 4, 7, 10, 13)`.
```{r, echo=FALSE, comment=NA}

```

If a variable `phone_number` contains the contact number of each person, what type of variable is `phone_number`?

Which of the following file extensions is used to save R script files?
   a) .doc  
   b) .csv  
   c) .R  
   d) .xlsx  

Why might the median be preferred over the mean when summarizing the central tendency of skewed data?

What function can be used to view the first few rows of a dataframe object? Implement it on an object called `data_frame_example`. (Note this code does not need to run, just show the correct function)

What are range, interquartile range (IQR), and standard deviation attempting to measure?

What is it called when a distribution has only one mode?

If a variable `education_level` is coded such that high school = 1, bachelor = 2, master = 3, and PhD = 4, what type of variable is `education_level`?

Name one statistic that is not resistant.

What is the main difference between a boxplot and a scatterplot?


---