### **Practice Exam: DATA 180-01**
---
title: "DATA 180-01 Exam Practice with Key"
subtitle: "Fall 2024, Professor Kessler"
author: 
  name: "Your Name"
  email: "XXXX@dickinson.edu"
date: "`r format(Sys.Date(), "%B %d, %Y")`"
output: 
  html_document
---

## Problem 1
(Each part worth 3 points)

Create a vector named `departments` with the names of different company departments: "Finance", "HR", "IT", "Marketing", "Operations"

```{r}
departments <- c("Finance", "HR", "IT", "Marketing", "Operations")
departments
```

a) The `Marketing` department has been renamed to `Sales and Marketing`. Update the vector to reflect this change.

```{r}
# Replace the incorrect entry in the vector
departments[4] <- "Sales and Marketing"
```

b) Write a function named `print_departments` that takes a vector of departments and prints each department along with its position in the vector (e.g., "1: Finance").

```{r}
# Create the function to print each department and its position
print_departments <- function(departments) {
  for (i in departments) {
    print(i)
  }
}

print_departments(departments)
```

c) Create a data frame named `department_budget` that contains the following information:

|Department         | Budget ($) in Millions |
|:------------------|:-----------------------|
|Finance            | 5.0                    |
|HR                 | 2.0                    |
|IT                 | 8.0                    |
|Sales and Marketing| 4.5                    |
|Operations         | 6.0                    |

```{r}
# Create the data frame
department_budget <- data.frame(
  Department = departments,
  Budget = c(5.0, 2.0, 8.0, 4.5, 6.0)
)
department_budget
```

d) The budget of the `IT` department has increased by 1.5 million. Update the `department_budget` data frame accordingly.

```{r}
# Update the budget of IT department
department_budget[department_budget$Department == "IT",]$Budget <- department_budget[department_budget$Department == "IT",]$Budget + 1.5
department_budget
```

e) Calculate the total budget of all departments.

```{r}
# Calculate the total budget
total_budget <- sum(department_budget$Budget)
total_budget
```
f) Calculate the mean and median budget.

```{r}
# Calculate the total budget
mean(department_budget$Budget)
median(department_budget$Budget)
```


## Problem 2
(Each part worth 4 points)

Consider the sales data of a company for two products, `Product_A` and `Product_B`, for each quarter of the year:

```{r}
sales_data <- data.frame(
  Quarter = c("Q1", "Q2", "Q3", "Q4"),
  Product_A = c(15000, 22000, 18000, 24000),
  Product_B = c(18000, 20000, 21000, 23000)
)
sales_data
```

a) Create a new column named `Total_Sales` that contains the total sales for both products in each quarter.

```{r}
# Create the Total_Sales column
sales_data$Total_Sales <- sales_data$Product_A + sales_data$Product_B
sales_data
```

b) Write a conditional statement that creates a new column named `Performance` with values "Good" if total sales exceed 40000, otherwise "Needs Improvement".

```{r}
# Create the Performance column
perf <- c()
counter <-1
for (i in sales_data$Total_Sales){
  if (i > 40000){
    perf[counter]<-"Good"
  }
  else{
    perf[counter]<-"Needs Improvement"
  }
  counter <- counter+1
}
sales_data$Performance <-perf
sales_data
```

c) Plot a bar chart that shows the total sales of `Product_A` and `Product_B` combined across all quarters. Do this in both base R and ggplot2.

```{r}
# Plot the bar chart
library(ggplot2)
barplot(sales_data$Total_Sales,col=c(1:4),names=sales_data$Quarter,xlab="Quarters",ylab="Total Sales")
ggplot(sales_data,aes(Quarter,Total_Sales))+geom_bar(stat="identity",fill=1:4)
```

d) Write a function named `quarter_summary` that takes the sales data and returns the average sales of `Product_A` and `Product_B`.

```{r}
# Create the function to calculate average sales
quarter_summary <- function(sales_data) {
  avg_product_a <- mean(sales_data$Product_A)
  avg_product_b <- mean(sales_data$Product_B)
  result <- c(avg_product_a,avg_product_b)
  return(result)
}

quarter_summary(sales_data)
```

---

## Problem 3
(Each part worth 5 points)

We have two data frames: `students` with student names and their majors, and `courses` with course details:

```{r}
students <- data.frame(
  Student = c("Alice", "Bob", "Charlie", "David", "Eve"),
  Major = c("Mathematics", "Physics", "Economics", "Computer Science", "Biology")
)

courses <- data.frame(
  Course = c("Calculus", "Physics 101", "Econometrics", "Programming 101", "Genetics"),
  Major = c("Mathematics", "Physics", "Economics", "Computer Science", "Biology")
)
```



a) Create a variable named `course_difficulty` for each course with levels: "Easy", "Medium", "Hard". Assign them as follows: `Calculus` - "Medium", `Physics 101` - "Easy", `Econometrics` - "Hard", `Programming 101` - "Easy", `Genetics` - "Medium".

```{r}
# Create the factor variable
course_difficulty <- c("Medium", "Easy", "Hard", "Easy", "Medium")
courses$difficulty<-course_difficulty
course_difficulty
```

b) Write a function named `course_summary` that takes a data frame of courses and returns the number of courses for each difficulty level.

```{r}
# Create the function to count courses by difficulty level
course_summary <- function(courses) {
  table(courses$difficulty)
}

course_summary(courses)
```



---

## Practice Questions

Can a variable be both continuous and discrete?  
- No, a variable cannot be both continuous and discrete. Discrete variables are countable and finite, whereas continuous variables can take any value within a given range. They represent different types of data measurements.

If a variable `weight_kg` contains a numeric value representing weight with up to two decimal places, what type of variable is `weight_kg`?
- The variable `weight_kg` is a continuous variable since it can take any value within a range and is measured to a specific level of precision.

What is the main purpose of the `==` operator in R?* 
- The `==` operator in R is used for **equality comparison. It checks if two values are equal and returns `TRUE` or `FALSE`.

Write a single line of R code that returns the following sequence: (1, 4, 7, 10, 13). Do not simply write out `c(1, 4, 7, 10, 13)`. 
seq(1, 13, by = 3)


If a variable `phone_number` contains the contact number of each person, what type of variable is `phone_number`?
- The variable `phone_number` is a nominal variable. Although it contains numeric values, it is used as an identifier rather than for quantitative measurement.

Which of the following file extensions is used to save R script files? 
   c) .R

Why might the median be preferred over the mean when summarizing the central tendency of skewed data?  
- The median is resistant to outliers, meaning it is less affected by extremely large or small values compared to the mean, making it a better measure of central tendency for skewed data.

What function can be used to view the first few rows of a dataframe object? Implement it on an object called `data_frame_example`. 
- The head() function is used to view the first few rows of a dataframe object.

What are range, interquartile range (IQR), and standard deviation attempting to measure?
- Range, IQR, and standard deviation are all measures of variability or dispersion. They help determine how spread out the data is around the center.

What is it called when a distribution has one mode? 
- A distribution with one mode is called unimodAl.

If a variable `education_level` is coded such that high school = 1, bachelor = 2, master = 3, and PhD = 4, what type of variable is `education_level`?  
- The variable `education_level` is an ordinal variable, as it represents categories with a natural order but no consistent interval between them.

Name one statistic that is not resistant.
- The mean is not resistant, as it is affected by outliers and extreme values.

What is the main difference between a boxplot and a scatterplot? 
- A boxplot is used to visualize the distribution of a single variable, showing the median, quartiles, and potential outliers. A scatterplot is used to show the relationship between two variables, often used to determine if a correlation exists.

--- 


