
---
title: "dplyr Practice Problems"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

This document contains a set of practice problems designed to help you become familiar with the `dplyr` package in R. The dataset we'll be using is the `diamonds` dataset from the `ggplot2` package.

```{r, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2) # Contains the diamonds dataset
```

# Problem 1: Exploring the Dataset

**Question:**

1. Load the `diamonds` dataset.
2. Display the first 6 rows of the dataset.
3. Use `glimpse()` to get an overview of the dataset structure.



```{r}
# 1. Load the diamonds dataset (it's already loaded with ggplot2)
data(diamonds)

# 2. Display the first 6 rows
head(diamonds)

# 3. Use glimpse to get an overview
glimpse(diamonds)
```

# Problem 2: Selecting Columns

**Question:**

1. Create a new data frame called `diamonds_small` that contains only the following columns: `carat`, `cut`, `color`, `price`.
2. Display the first 6 rows of `diamonds_small`.



```{r}
# 1. Select specified columns
diamonds_small <- diamonds %>%
  select(carat, cut, color, price)

# 2. Display first 6 rows
head(diamonds_small)
```

# Problem 3: Filtering Rows

**Question:**

1. Filter the `diamonds` dataset to include only diamonds with a `cut` of "Ideal" and a `color` of "E".
2. Assign this filtered dataset to a new object called `ideal_E_diamonds`.
3. How many diamonds meet these criteria?



```{r}
# 1 & 2. Filter and assign
ideal_E_diamonds <- diamonds %>%
  filter(cut == "Ideal", color == "E")

# 3. Count the number of diamonds
nrow(ideal_E_diamonds)
```

# Problem 4: Arranging Rows

**Question:**

1. Arrange the `diamonds` dataset in descending order of `price`.
2. Display the first 10 rows of the arranged dataset.



```{r}
# 1. Arrange in descending order of price
diamonds_desc <- diamonds %>%
  arrange(desc(price))

# 2. Display first 10 rows
head(diamonds_desc, 10)
```

# Problem 5: Creating New Variables

**Question:**

1. Create a new variable in the `diamonds` dataset called `price_per_carat` that is calculated as `price / carat`.
2. Display the first 6 rows of the modified dataset, showing the `carat`, `price`, and `price_per_carat` columns.



```{r}
# 1. Create new variable
diamonds <- diamonds %>%
  mutate(price_per_carat = price / carat)

# 2. Display selected columns
diamonds %>%
  select(carat, price, price_per_carat) %>%
  head()
```

# Problem 6: Summarizing Data

**Question:**

1. Calculate the average `price` of diamonds for each `cut` type.
2. Display the results in a data frame with two columns: `cut` and `average_price`.



```{r}
# 1. Summarize data
average_price_by_cut <- diamonds %>%
  group_by(cut) %>%
  summarise(average_price = mean(price))

# 2. Display results
average_price_by_cut
```

# Problem 7: Chaining Multiple Operations

**Question:**

1. From the `diamonds` dataset, select diamonds with a `carat` greater than 1.5.
2. For these diamonds, create a new variable `price_per_carat` as before.
3. Arrange the data in ascending order of `price_per_carat`.
4. Select only the `carat`, `cut`, `color`, `price_per_carat` columns.
5. Display the first 10 rows of the resulting dataset.



```{r}
diamonds %>%
  filter(carat > 1.5) %>%
  mutate(price_per_carat = price / carat) %>%
  arrange(price_per_carat) %>%
  select(carat, cut, color, price_per_carat) %>%
  head(10)
```

# Problem 8: Counting and Tallying

**Question:**

1. Count the number of diamonds for each combination of `cut` and `color`.
2. Arrange the results in descending order of the count.
3. Display the top 10 combinations.



```{r}
# 1 & 2. Count and arrange
diamond_counts <- diamonds %>%
  count(cut, color) %>%
  arrange(desc(n))

# 3. Display top 10 combinations
head(diamond_counts, 10)
```

# Problem 9: Sampling Data

**Question:**

1. Take a random sample of 100 diamonds from the `diamonds` dataset.
2. Display the `carat`, `cut`, `color`, and `price` of the sampled diamonds.



```{r}
set.seed(123) # For reproducibility

sampled_diamonds <- diamonds %>%
  sample_n(100) %>%
  select(carat, cut, color, price)

# Display the sampled data
sampled_diamonds
```



---
