
---
title: "dplyr Practice Problems"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

This document contains a set of practice problems designed to help you become familiar with the `dplyr` package in R. The dataset we'll be using is the `diamonds` dataset from the `ggplot2` package.

```{r, include=FALSE}
# Load necessary packages
library(dplyr)
library(ggplot2) # Contains the diamonds dataset
```

# Problem 1: Exploring the Dataset

**Question:**

1. Load the `diamonds` dataset.
2. Display the first 6 rows of the dataset.
3. Use `glimpse()` to get an overview of the dataset structure.



```{r}

```

# Problem 2: Selecting Columns

**Question:**

1. Create a new data frame called `diamonds_small` that contains only the following columns: `carat`, `cut`, `color`, `price`.
2. Display the first 6 rows of `diamonds_small`.



```{r}

```

# Problem 3: Filtering Rows

**Question:**

1. Filter the `diamonds` dataset to include only diamonds with a `cut` of "Ideal" and a `color` of "E".
2. Assign this filtered dataset to a new object called `ideal_E_diamonds`.
3. How many diamonds meet these criteria?



```{r}


```

# Problem 4: Arranging Rows

**Question:**

1. Arrange the `diamonds` dataset in descending order of `price`.
2. Display the first 10 rows of the arranged dataset.



```{r}

```

# Problem 5: Creating New Variables

**Question:**

1. Create a new variable in the `diamonds` dataset called `price_per_carat` that is calculated as `price / carat`.
2. Display the first 6 rows of the modified dataset, showing the `carat`, `price`, and `price_per_carat` columns.



```{r}

```

# Problem 6: Summarizing Data

**Question:**

1. Calculate the average `price` of diamonds for each `cut` type.
2. Display the results in a data frame with two columns: `cut` and `average_price`.



```{r}

```

# Problem 7: Chaining Multiple Operations

**Question:**

1. From the `diamonds` dataset, select diamonds with a `carat` greater than 1.5.
2. For these diamonds, create a new variable `price_per_carat` as before.
3. Arrange the data in ascending order of `price_per_carat`.
4. Select only the `carat`, `cut`, `color`, `price_per_carat` columns.
5. Display the first 10 rows of the resulting dataset.



```{r}

```

# Problem 8: Counting and Tallying

**Question:**

1. Count the number of diamonds for each combination of `cut` and `color`.
2. Arrange the results in descending order of the count.
3. Display the top 10 combinations.



```{r}

```

# Problem 9: Sampling Data

**Question:**

1. Take a random sample of 100 diamonds from the `diamonds` dataset.
2. Display the `carat`, `cut`, `color`, and `price` of the sampled diamonds.



```{r}

```



---
