---
title: "DATA 180 Final Exam Exercises"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
#install.packages("gapminder")
library(gapminder)
library(tidyverse)
library(cluster)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 5,
  fig.height = 4
)
```

## Question 1
Using the `mtcars` dataset, identify key features of fuel-efficient (mpg>25) cars and compare their similarities using a distance measure. How similar are they on average? What is the average distance between fuel-efficient and inefficient vehicles?
```{r, echo=TRUE}

```

## Question 2
In the `iris` dataset, examine the relationship between sepal length and petal length through a scatterplot. Describe any noticeable patterns. 
```{r, echo=TRUE}

```

## Question 3
Evaluate how the cut quality of diamonds affects their average price in the `diamonds` dataset. Use a bar chart to represent your findings.
```{r, echo=TRUE}

```

## Question 4
Using the `USArrests` dataset, apply hierarchical clustering to normalized data. Represent the clusters in a dendrogram and describe how the clusters change with different methods.
```{r, echo=TRUE}

```

## Question 5
Cluster observations in the `iris` dataset based on petal dimensions. Visualize the clusters using a suitable scatterplot.
```{r, echo=TRUE}

```

## Question 6
Assess vehicle similarities in the `mtcars` dataset by calculating a Gower distance matrix. Perform clustering and interpret the groups.
```{r, echo=TRUE}

```

## Question 7
Perform a regression analysis in the `mtcars` dataset to determine how weight and horsepower affect fuel efficiency. Which impacts fuel efficiency more?
```{r, echo=TRUE}

```

## Question 8
Analyze GDP per capita trends over time in the `gapminder` dataset. Highlight the changes across decades and provide a summary.
```{r, echo=TRUE}

```

## Question 9
Using the `storms` dataset, analyze variations in maximum wind speed across storm categories. Create a line or bar graph to visualize your findings.
```{r, echo=TRUE}

```

## Question 10
Explore the relationship between horsepower and weight in the `mtcars` dataset. Use a scatterplot to summarize your findings and add a regression line.
```{r, echo=TRUE}

```

## Question 11
Investigate the correlation between sepal dimensions for each species in the `iris` dataset. Summarize the results with visuals and descriptive text.
```{r, echo=TRUE}

```

## Question 12
Apply hierarchical clustering to the `USArrests` dataset and identify key patterns. Use a dendrogram to visualize the clusters.
```{r, echo=TRUE}

```

## Question 13
Perform a multiple regression analysis in the `mtcars` dataset to predict fuel efficiency. Identify the most (statistically) significant predictors.
```{r, echo=TRUE}

```

## Question 14
Explore GDP per capita trends in Asia over time using the `gapminder` dataset. Only look at observations after 1990 and countries with a population above 1 million at that point. Compare and contrast the trajectories for different countries. What is the relationship between GDP per capita and life expectancy?
```{r, echo=TRUE}

```

## Question 15
Use clustering techniques on the `USArrests` dataset to group states into clusters and visualize the clusters using a scatterplot of `Murder` vs `Assault`. Be sure to choose the optimal number of clusters.
```{r, echo=TRUE}
set.seed(42)

```
