
---
title: "Practice Questions for Exam 2"
output: html_document
---

```{r, include=FALSE}
library(dplyr)     
library(ggplot2)    
    
```




### Question 1
Load the `mtcars` dataset. Use `dplyr` to filter rows where the `mpg` is greater than 25 and display the first 5 rows.

```{r}
data(mtcars)
```

### Question 2
Using the `iris` dataset, group the data by `Species` and calculate the mean `Sepal.Length` for each group.

```{r}
data(iris)
```

### Question 3
Select only the `cyl` and `hp` columns from the `mtcars` dataset and arrange the rows in descending order of `hp`.

```{r}
data(mtcars)
```

### Question 4
Use `mutate` on the `mtcars` dataset to create a new column `weight_kg` by converting the `wt` column (in 1000 lbs) into kilograms.

```{r}
data(mtcars)
```

### Question 5
Filter the `diamonds` dataset to show only diamonds with a `carat` greater than 2.5 and a `price` less than 10,000.

```{r}
data(diamonds)
```

### Question 6
Write a pipeline to:
- Filter rows in `iris` where `Sepal.Width` > 3.
- Select the columns `Sepal.Length` and `Species`.
- Arrange the resulting rows by `Sepal.Length` in ascending order.

```{r}
data(iris)
```

### Question 7
Summarize the `mtcars` dataset to find the average horsepower (`hp`) for each unique number of cylinders (`cyl`).

```{r}
data(mtcars)
```

### Question 8
Add a new column to the `iris` dataset that categorizes `Petal.Length` as "Short" (< 2), "Medium" (2-4), or "Long" (> 4). Hint: use the case_when function which lets you specific that when a value in a column meets a certain condition, it takes on a certain value. Ex. case_when(Petal.Length == 2 ~ "Short")

```{r}
data(iris)
```


### Question 9
Compute the Euclidean distances between rows of the `mtcars` dataset using only the columns `mpg`, `disp`, and `hp`.

```{r}
data(mtcars)
```

### Question 10
Using the `iris` dataset, calculate the Manhattan distances between the first 10 rows using `Sepal.Length` and `Petal.Width`.

```{r}
data(iris)
```

### Question 11
Using the `mtcars` dataset's `mpg` and `hp` columns compute the Euclidean distances.

```{r}
data(mtcars)
```


### Question 12
Perform hierarchical clustering on the `iris` dataset using only the `Sepal.Length` and `Sepal.Width` columns. Use the `complete` linkage method and plot the resulting dendrogram.

```{r}
data(iris)
```

### Question 13
Using the `mtcars` dataset, cluster the cars based on the `mpg`, `hp`, and `wt` columns. Use the average linkage method.

```{r}
data(mtcars)
```

### Question 14
Write R code to put rectangles around 3 clusters in the dendrogram from clustering the `iris` dataset.

```{r}
data(iris)
```

### Question 15
Write R code to extract the cluster assignments for 3 clusters from a dendrogram created using the `iris` dataset. Hint: Use cuttree.

```{r}
data(iris)
```

### Question 16
Perform k-means clustering on the `iris` dataset using `Petal.Length` and `Petal.Width`. Use `k = 3` and summarize the number of data points in each cluster.

```{r}
data(iris)
```

### Question 17
Using the `mtcars` dataset, determine the optimal number of clusters (`k`) using the Elbow Method for columns `mpg`, `hp`, and `wt`.

```{r}
data(mtcars)
```

### Question 18
Plot the k-means clusters from the `iris` dataset (using `Petal.Length` and `Petal.Width`) on a scatterplot, coloring the points by their cluster assignments.

```{r}
data(iris)
```

### Question 19
Explain how the initial random centroids in k-means can impact the final cluster assignments. Perform k-means clustering on the `mtcars` dataset twice. Only use the mpg, hp, and wt columns.

```{r}
data(mtcars)

set.seed(123)
```

### Question 20
Compare the performance of k-means clustering and hierarchical clustering on the `mtcars` dataset based on columns `mpg`, `hp`, and `wt`. Do they broadly align in the placing the same items in the same clusters?

```{r}
data(mtcars)

```
