---

title: "Practice Exercises for Exam 1"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document

---

## Exercise: Assigning Variables in R

1. Assign the value `42` to a variable named `my_number`.
2. Assign the string `"Hello, R!"` to a variable named `greeting`.
3. Assign the logical value `FALSE` to a variable named `is_sunny`.

```{r}
# Your code here

```

---

## Exercise: Basic Operations in R

1. Calculate the sum of `25` and `17`.
2. Multiply `6` by `7` and assign the result to a variable named `product`.
3. Divide `100` by `4` and subtract `15` from the result.

```{r}
# Your code here

```

---

## Exercise: Creating Vectors

1. Create a numeric vector named `temperatures` containing the values: `23, 17, 21, 19, 25`.
2. Create a character vector named `fruits` containing the strings: `"apple"`, `"banana"`, `"cherry"`.
3. Create a sequence of numbers from `1` to `20` and assign it to a variable named `my_sequence`.

```{r}
# Your code here

```

---

## Exercise: Basic Statistics

Using the `temperatures` vector from the previous exercise:

1. Calculate the mean temperature.
2. Find the median temperature.
3. Calculate the variance and standard deviation of the temperatures.

```{r}
# Your code here

```

---

## Exercise: Writing For Loops

1. Write a for loop that prints the numbers from `1` to `10`.
2. Using the `fruits` vector, write a for loop that prints each fruit in the vector.
3. Write a for loop that calculates the squares of numbers from `1` to `5` and stores them in a vector named `squares`.

```{r}
# Your code here

```

---

## Exercise: Creating Simple Functions

1. Write a function named `add_three` that takes a number as input and returns the number plus three.
2. Create a function named `is_even` that returns `TRUE` if a number is even and `FALSE` if it's odd.
3. Define a function named `fahrenheit_to_celsius` that converts a temperature from Fahrenheit to Celsius.

```{r}
# Your code here

```

---

## Exercise: Creating Simple Functions

1. Write a function named `add_three` that takes a number as input and returns the number plus three.
2. Create a function named `is_even` that returns `TRUE` if a number is even and `FALSE` if it's odd.
3. Define a function named `fahrenheit_to_celsius` that converts a temperature from Fahrenheit to Celsius.

```{r}
# Solution

# 1. Function to add three
add_three <- function(x) {
  return(x + 3)
}

# 2. Function to check if a number is even
is_even <- function(x) {
  return(x %% 2 == 0)
}

# 3. Function to convert Fahrenheit to Celsius
fahrenheit_to_celsius <- function(f) {
  c <- (f - 32) * 5/9
  return(c)
}
```

---

## Exercise: Creating Data Frames

1. Create a data frame named `students` with the following columns and data:

   | Name     | Age | Major            |
   |----------|-----|------------------|
   | Alice    | 20  | Mathematics      |
   | Bob      | 22  | Physics          |
   | Charlie  | 19  | Engineering      |
   | Diana    | 21  | Biology          |
   | Ethan    | 23  | Computer Science |
   | Fiona    | 20  | Chemistry        |
   | George   | 22  | Economics        |
   | Hannah   | 21  | Psychology       |
   | Irene    | 20  | Art              |
   | Jack     | 22  | Literature       |

2. Add a new column to the `students` data frame named `GPA` with the values `3.5`, `3.8`, `3.2`, `3.7`, `3.9`, `3.4`, `3.6`, `3.8`, `3.3`, and `3.7`.

3. Access the `Major` column from the `students` data frame.

```{r}
# Your code here

```

---

## Exercise: Importing Data Frames from Packages

1. Install and load the `MASS` package.
2. Load the `Boston` data set from the `MASS` package.
3. Display the first six rows of the `Boston` data set.

```{r}
# Your code here

```

---

## Exercise: Basic Plots with Base R

Using the `Boston` data set:

1. Create a scatter plot of `medv` (median value of owner-occupied homes) vs. `lstat` (percentage of lower status population).
2. Create a histogram of the `crim` (per capita crime rate by town) variable.
3. Create a bar plot of the `chas` (Charles River dummy variable) variable.

```{r}
# Your code here

```

---

## Exercise: Barplots and Pie Charts with Base R

1. Create a barplot of the number of students in each `Major` from the `students` data frame.
2. Create a pie chart of the distribution of `Major` in the `students` data frame.

```{r}
# Your code here

```

---

## Exercise: Scatter Plots with Base R

Using the `cars` dataset:

1. Create a scatter plot of `speed` vs. `dist`.
2. Add a title and axis labels to the scatter plot.

```{r}
# Your code here

```

---

## Exercise: Data Visualization with ggplot2

1. Install and load the `ggplot2` package.
2. Using the `mpg` dataset, create a scatter plot of `displ` vs. `hwy`.
3. Create a bar plot showing the count of cars for each `class`.
4. Create a histogram of the `cty` (city miles per gallon) variable.

```{r}
# Your code here

```

---

## Exercise: Advanced ggplot2 Visualizations

Using the `mpg` dataset:

1. Create a scatter plot of `displ` vs. `hwy` colored by `class`.
2. Add a smoothing line (`geom_smooth`) to the scatter plot.
3. Facet the scatter plot by `year`.

```{r}
# Your code here

```

---

## Exercise: Descriptive Statistics

Using the `iris` dataset:

1. Calculate the mean, median, variance, and standard deviation for the `Sepal.Length` variable.
2. Find the mode of the `Species` variable.
3. Summarize the `iris` dataset using the `summary()` function.

```{r}
# Your code here

```

---

## Exercise: Data Manipulation

1. Using the `dplyr` package, select the `Sepal.Length` and `Species` columns from the `iris` dataset.
2. Filter the `iris` dataset for rows where `Species` is `"setosa"`.
3. Arrange the `iris` dataset in descending order of `Petal.Width`.

```{r}
# Your code here

```

---

## Exercise: Calculating Frequencies

1. Using the `table()` function, calculate the frequency of each `Species` in the `iris` dataset.
2. Create a bar plot of the species frequencies.

```{r}
# Your code here

```


---

## Exercise: Control Structures

1. Write an `if` statement that checks if a number is positive, negative, or zero.
2. Use a `while` loop to calculate the factorial of `5`.

```{r}
# Your code here

```

