---

title: "Practice Exercises for Exam 1"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---


## Exercise: Assigning Variables in R

1. Assign the value `42` to a variable named `my_number`.
2. Assign the string `"Hello, R!"` to a variable named `greeting`.
3. Assign the logical value `FALSE` to a variable named `is_sunny`.

```{r}
# Solution

my_number <- 42
greeting <- "Hello, R!"
is_sunny <- FALSE
```

---

## Exercise: Basic Operations in R

1. Calculate the sum of `25` and `17`.
2. Multiply `6` by `7` and assign the result to a variable named `product`.
3. Divide `100` by `4` and subtract `15` from the result.

```{r}
# Solution

# 1. Sum
sum_result <- 25 + 17
sum_result  # Output: 42

# 2. Multiplication
product <- 6 * 7
product  # Output: 42

# 3. Division and Subtraction
result <- (100 / 4) - 15
result  # Output: 10
```

---

## Exercise: Creating Vectors

1. Create a numeric vector named `temperatures` containing the values: `23, 17, 21, 19, 25`.
2. Create a character vector named `fruits` containing the strings: `"apple"`, `"banana"`, `"cherry"`.
3. Create a sequence of numbers from `1` to `20` and assign it to a variable named `my_sequence`.

```{r}
# Solution

# 1. Numeric vector
temperatures <- c(23, 17, 21, 19, 25)

# 2. Character vector
fruits <- c("apple", "banana", "cherry")

# 3. Sequence
my_sequence <- 1:20
```

---

## Exercise: Basic Statistics

Using the `temperatures` vector from the previous exercise:

1. Calculate the mean temperature.
2. Find the median temperature.
3. Calculate the variance and standard deviation of the temperatures.

```{r}
# Solution

# 1. Mean
mean_temp <- mean(temperatures)
mean_temp  # Output: 21

# 2. Median
median_temp <- median(temperatures)
median_temp  # Output: 21

# 3. Variance and Standard Deviation
variance_temp <- var(temperatures)
sd_temp <- sd(temperatures)
variance_temp  # Output: 10
sd_temp        # Output: 3.162278
```

---

## Exercise: Writing For Loops

1. Write a for loop that prints the numbers from `1` to `10`.
2. Using the `fruits` vector, write a for loop that prints each fruit in the vector.
3. Write a for loop that calculates the squares of numbers from `1` to `5` and stores them in a vector named `squares`.

```{r}
# Solution

# 1. Printing numbers from 1 to 10
for (i in 1:10) {
  print(i)
}

# 2. Printing each fruit
for (fruit in fruits) {
  print(fruit)
}

# 3. Calculating squares
squares <- c()
for (i in 1:5) {
  squares[i] <- i^2
}
squares  # Output: 1 4 9 16 25
```

---

## Exercise: Creating Simple Functions

1. Write a function named `add_three` that takes a number as input and returns the number plus three.
2. Create a function named `is_even` that returns `TRUE` if a number is even and `FALSE` if it's odd.
3. Define a function named `fahrenheit_to_celsius` that converts a temperature from Fahrenheit to Celsius.

```{r}
# Solution

# 1. Function to add three
add_three <- function(x) {
  return(x + 3)
}

# 2. Function to check if a number is even
is_even <- function(x) {
  return(x %% 2 == 0)
}

# 3. Function to convert Fahrenheit to Celsius
fahrenheit_to_celsius <- function(f) {
  c <- (f - 32) * 5/9
  return(c)
}
```

---

## Exercise: Creating Data Frames

1. Create a data frame named `students` with the following columns and data:

   | Name     | Age | Major            |
   |----------|-----|------------------|
   | Alice    | 20  | Mathematics      |
   | Bob      | 22  | Physics          |
   | Charlie  | 19  | Data Analytics   |
   | Diana    | 21  | Biology          |
   | Ethan    | 23  | Computer Science |
   | Fiona    | 20  | Data Analytics   |
   | George   | 22  | Data Analytics   |
   | Hannah   | 21  | Psychology       |
   | Irene    | 20  | Psychology       |
   | Jack     | 22  | Physics          |

2. Add a new column to the `students` data frame named `GPA` with the values `3.5`, `3.8`, `3.2`, `3.7`, `3.9`, `3.4`, `3.6`, `3.8`, `3.3`, and `3.7`.

3. Access the `Major` column from the `students` data frame.

```{r}
# Solution

# 1. Creating the data frame
students <- data.frame(
  Name = c("Alice", "Bob", "Charlie", "Diana", "Ethan", "Fiona", "George", "Hannah", "Irene", "Jack"),
  Age = c(20, 22, 19, 21, 23, 20, 22, 21, 20, 22),
  Major = c("Mathematics", "Physics", "Data Analytics", "Biology", "Computer Science", "Data Analytics", "Data Analytics", "Psychology", "Psychology", "Physics")
)

# 2. Adding the GPA column
students$GPA <- c(3.5, 3.8, 3.2, 3.7, 3.9, 3.4, 3.6, 3.8, 3.3, 3.7)

# 3. Accessing the Major column
majors <- students$Major
majors  # Output: vector of majors
```

---

## Exercise: Importing Data Frames from Packages

1. Install and load the `MASS` package.
2. Load the `Boston` data set from the `MASS` package.
3. Display the first six rows of the `Boston` data set.

```{r}
# Solution

# 1. Install and load MASS package
# install.packages("MASS")
library(MASS)

# 2. Load the Boston dataset
data(Boston)

# 3. Display the first six rows
head(Boston)
```

---

## Exercise: Basic Plots with Base R

Using the `Boston` data set:

1. Create a scatter plot of `medv` (median value of owner-occupied homes) vs. `lstat` (percentage of lower status population).
2. Create a histogram of the `crim` (per capita crime rate by town) variable.
3. Create a bar plot of the `chas` (Charles River dummy variable) variable.

```{r}
# Solution

# 1. Scatter plot
plot(Boston$lstat, Boston$medv,
     xlab = "LSTAT (%)",
     ylab = "MEDV ($1000s)",
     main = "MEDV vs LSTAT",
     pch = 19,
     col = "blue")

# 2. Histogram
hist(Boston$crim,
     breaks = 30,
     main = "Histogram of CRIM",
     xlab = "Per Capita Crime Rate",
     col = "lightblue")

# 3. Bar plot
chas_counts <- table(Boston$chas)
barplot(chas_counts,
        main = "Counts of CHAS Variable",
        xlab = "CHAS",
        ylab = "Frequency",
        col = c("orange", "purple"))
```

---

## Exercise: Barplots and Pie Charts with Base R

1. Create a barplot of the number of students in each `Major` from the `students` data frame.
2. Create a pie chart of the distribution of `Major` in the `students` data frame.

```{r}
# Solution

# 1. Barplot
major_counts <- table(students$Major)
barplot(major_counts/nrow(students),
        main = "Number of Students per Major",
        xlab = "Major",
        ylab = "Number of Students",
        col = "skyblue")

# 2. Pie chart
pie(major_counts,
    main = "Distribution of Majors",
    labels = names(major_counts),
    col = rainbow(length(major_counts)))
```

---

## Exercise: Scatter Plots with Base R

Using the `cars` dataset:

1. Create a scatter plot of `speed` vs. `dist`.
2. Add a title and axis labels to the scatter plot.

```{r}
# Solution

plot(cars$speed, cars$dist,
     xlab = "Speed (mph)",
     ylab = "Stopping Distance (ft)",
     main = "Stopping Distance vs Speed",
     pch = 19,
     col = "darkgreen")
```

---

## Exercise: Data Visualization with ggplot2

1. Install and load the `ggplot2` package.
2. Using the `mpg` dataset, create a scatter plot of `displ` vs. `hwy`.
3. Create a bar plot showing the count of cars for each `class`.
4. Create a histogram of the `cty` (city miles per gallon) variable.

```{r}
# Solution

# 1. Install and load ggplot2
# install.packages("ggplot2")  # Uncomment if not already installed
library(ggplot2)

# 2. Scatter plot
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  labs(title = "Engine Displacement vs Highway MPG",
       x = "Displacement (L)",
       y = "Highway MPG")

# 3. Bar plot
ggplot(data = mpg, aes(x = class)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Count of Cars by Class",
       x = "Car Class",
       y = "Count")

# 4. Histogram
ggplot(data = mpg, aes(x = cty)) +
  geom_histogram(binwidth = 2, fill = "coral", color = "black") +
  labs(title = "Histogram of City MPG",
       x = "City MPG",
       y = "Frequency")
```

---

## Exercise: Advanced ggplot2 Visualizations

Using the `mpg` dataset:

1. Create a scatter plot of `displ` vs. `hwy` colored by `class`.
2. Add a smoothing line (`geom_smooth`) to the scatter plot.
3. Facet the scatter plot by `year`.

```{r}
# Solution

# 1. Scatter plot colored by class
ggplot(data = mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  labs(title = "Displacement vs Highway MPG by Class",
       x = "Displacement (L)",
       y = "Highway MPG")

# 2. Add smoothing line
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(title = "Displacement vs Highway MPG with Smoothing Line",
       x = "Displacement (L)",
       y = "Highway MPG")

# 3. Facet by year
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_wrap(~ year) +
  labs(title = "Displacement vs Highway MPG Faceted by Year",
       x = "Displacement (L)",
       y = "Highway MPG")
```

---

## Exercise: Descriptive Statistics

Using the `iris` dataset:

1. Calculate the mean, median, variance, and standard deviation for the `Sepal.Length` variable.
2. Find the mode of the `Species` variable.
3. Summarize the `iris` dataset using the `summary()` function.

```{r}
# Solution

# 1. Descriptive statistics
mean_sepal <- mean(iris$Sepal.Length)
median_sepal <- median(iris$Sepal.Length)
variance_sepal <- var(iris$Sepal.Length)
sd_sepal <- sd(iris$Sepal.Length)

mean_sepal      # Output: 5.843333
median_sepal    # Output: 5.8
variance_sepal  # Output: 0.6856935
sd_sepal        # Output: 0.8280661

# 2. Mode of Species
species_counts <- table(iris$Species)
mode_species <- names(species_counts)[which.max(species_counts)]
mode_species   # Output: All species have equal counts

# 3. Summary of iris dataset
summary(iris)
```

---

## Exercise: Data Manipulation

1. Using the `dplyr` package, select the `Sepal.Length` and `Species` columns from the `iris` dataset.
2. Filter the `iris` dataset for rows where `Species` is `"setosa"`.
3. Arrange the `iris` dataset in descending order of `Petal.Width`.

```{r}
# Solution

# Load dplyr package
# install.packages("dplyr")  # Uncomment if not already installed
library(dplyr)

# 1. Select columns
selected_data <- select(iris, Sepal.Length, Species)
head(selected_data)

# 2. Filter for setosa species
setosa_data <- filter(iris, Species == "setosa")
head(setosa_data)

# 3. Arrange in descending order
arranged_data <- arrange(iris, desc(Petal.Width))
head(arranged_data)
```

---

## Exercise: Calculating Frequencies

1. Using the `table()` function, calculate the frequency of each `Species` in the `iris` dataset.
2. Create a bar plot of the species frequencies.

```{r}
# Solution

# 1. Calculating frequencies
species_freq <- table(iris$Species)
species_freq

# 2. Bar plot
barplot(species_freq,
        main = "Frequency of Iris Species",
        xlab = "Species",
        ylab = "Count",
        col = "lightgreen")
```

---


## Exercise: Control Structures

1. Write an `if` statement that checks if a number is positive, negative, or zero.
2. Use a `while` loop to calculate the factorial of `5`.

```{r}
# Solution

# 1. If statement
num <- -3
if (num > 0) {
  print("Positive")
} else if (num < 0) {
  print("Negative")
} else {
  print("Zero")
}

# 2. While loop for factorial
factorial_result <- 1
counter <- 5
while (counter > 0) {
  factorial_result <- factorial_result * counter
  counter <- counter - 1
}
factorial_result  # Output: 120
```

---