---
title: "Practice Questions for Exam 2"
output: html_document
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(cluster)
library(stats)
```

### Question 1
Load the `mtcars` dataset. Use `dplyr` to filter rows where the `mpg` is greater than 25 and display the first 5 rows.

```{r}
data(mtcars)
mtcars %>% filter(mpg > 25) %>% head(5)
```

### Question 2
Using the `iris` dataset, group the data by `Species` and calculate the mean `Sepal.Length` for each group.

```{r}
data(iris)
iris %>% group_by(Species) %>% summarize(mean_sepal_length = mean(Sepal.Length))
```

### Question 3
Select only the `cyl` and `hp` columns from the `mtcars` dataset and arrange the rows in descending order of `hp`.

```{r}
data(mtcars)
mtcars %>% select(cyl, hp) %>% arrange(desc(hp))
```

### Question 4
Use `mutate` on the `mtcars` dataset to create a new column `weight_kg` by converting the `wt` column (in 1000 lbs) into kilograms.

```{r}
data(mtcars)
mtcars %>% mutate(weight_kg = wt * 453.592)
```

### Question 5
Filter the `diamonds` dataset to show only diamonds with a `carat` greater than 2.5 and a `price` less than 10,000.

```{r}
data(diamonds, package = "ggplot2")
diamonds %>% filter(carat > 2.5, price < 10000)
```

### Question 6
Write a pipeline to:
- Filter rows in `iris` where `Sepal.Width` > 3.
- Select the columns `Sepal.Length` and `Species`.
- Arrange the resulting rows by `Sepal.Length` in ascending order.

```{r}
data(iris)
iris %>% filter(Sepal.Width > 3) %>% select(Sepal.Length, Species) %>% arrange(Sepal.Length)
```

### Question 7
Summarize the `mtcars` dataset to find the average horsepower (`hp`) for each unique number of cylinders (`cyl`).

```{r}
data(mtcars)
mtcars %>% group_by(cyl) %>% summarize(avg_hp = mean(hp))
```

### Question 8
Add a new column to the `iris` dataset that categorizes `Petal.Length` as "Short" (< 2), "Medium" (2-4), or "Long" (> 4).Hint: use the case_when function which lets you specific that when a value in a column meets a certain condition, it takes on a certain value. Ex. case_when(Petal.Length == 2 ~ "Short")

```{r}
data(iris)
iris %>% mutate(petal_length_category = case_when(
  Petal.Length < 2 ~ "Short",
  Petal.Length >= 2 & Petal.Length <= 4 ~ "Medium",
  Petal.Length > 4 ~ "Long"
))
```

### Question 9
Compute the pairwise Euclidean distances between rows of the `mtcars` dataset using only the columns `mpg`, `disp`, and `hp`.

```{r}
data(mtcars)
dist(mtcars %>% select(mpg, disp, hp))
```

### Question 10
Using the `iris` dataset, calculate the Manhattan distances between the first 10 rows using `Sepal.Length` and `Petal.Width`.

```{r}
data(iris)
dist(iris %>% slice(1:10) %>% select(Sepal.Length, Petal.Width), method = "manhattan")
```

### Question 11
Using the `mtcars` dataset's `mpg` and `hp` columns compute the Minkowski distances, p=1.5
```{r}
data(mtcars)
mtcars %>% select(mpg, hp) %>%  dist(method="minkowski",p=1.5)
```

### Question 12
Perform hierarchical clustering on the `iris` dataset using only the `Sepal.Length` and `Sepal.Width` columns. Use the `complete` linkage method and plot the resulting dendrogram.

```{r}
data(iris)
hc <- hclust(dist(iris %>% select(Sepal.Length, Sepal.Width)), method = "complete")
plot(hc)
```

### Question 13
Using the `mtcars` dataset, cluster the cars based on the `mpg`, `hp`, and `wt` columns. Use the `average` linkage method.

```{r}
data(mtcars)
hc <- hclust(dist(mtcars %>% select(mpg, hp, wt)), method = "average")
plot(hc)
```

### Question 14
Write R code to put rectangles around 3 clusters in the dendrogram from clustering the `iris` dataset.

```{r}
data(mtcars)
hc <- hclust(dist(mtcars %>% select(mpg, hp, wt)))
plot(hc)
rect.hclust(hc, k = 3, border = "red")
```

### Question 15
Write R code to extract the cluster assignments for 3 clusters from a dendrogram created using the `iris` dataset. Hint: Use cuttree.

```{r}
data(iris)
hc <- hclust(dist(iris %>% select(Sepal.Length, Sepal.Width)))
cutree(hc, k = 3)
```

### Question 16
Perform k-means clustering on the `iris` dataset using `Petal.Length` and `Petal.Width`. Use `k = 3` and summarize the number of data points in each cluster.

```{r}
data(iris)
kmeans_result <- kmeans(iris %>% select(Petal.Length, Petal.Width), centers = 3)
iris$clusters <- kmeans_result$cluster
iris %>% group_by(clusters) %>% summarize(mean_PL = mean(Petal.Length),mean_PW = mean(Petal.Width))
```

### Question 17
Using the `mtcars` dataset, determine the optimal number of clusters (`k`) using the Elbow Method for columns `mpg`, `hp`, and `wt`.

```{r}
data(mtcars)
wss <- c()
for (i in 1:10) {
  wss[i]<-kmeans(mtcars %>% select(mpg, hp, wt), centers = i)$tot.withinss
}
plot(1:10, wss, type = "b")
```

### Question 18
Plot the k-means clusters from the `iris` dataset (using `Petal.Length` and `Petal.Width`) on a scatterplot, coloring the points by their cluster assignments.

```{r}
data(iris)
kmeans_result <- kmeans(iris %>% select(Petal.Length, Petal.Width), centers = 3)
plot(iris$Petal.Length, iris$Petal.Width, col = kmeans_result$cluster, pch = 19, xlab = "Petal Length", ylab = "Petal Width")
```

### Question 19
Explain how the initial random centroids in k-means can impact the final cluster assignments. Perform k-means clustering on the `mtcars` dataset twice with different random seeds. Only use the mpg, hp, and wt columns.

```{r}
data(mtcars)
#set.seed(42)
kmeans_1 <- kmeans(mtcars %>% select(mpg, hp, wt), centers = 3)
set.seed(123)
kmeans_2 <- kmeans(mtcars %>% select(mpg, hp, wt), centers = 3)

plot(mpg~hp,data=mtcars,cex.axis=1.3,
     cex.lab=1.2,cex=1.2)
points(mpg~hp,data=kmeans_1$centers,pch=10,cex=1.8,col="orange")
points(mpg~hp,data=kmeans_2$centers,pch=10,cex=1.8,col="purple")
```

### Question 20
Compare the performance of k-means clustering and hierarchical clustering on the `mtcars` dataset based on columns `mpg`, `hp`, and `wt`. Do they broadly align in the placing the same items in the same clusters?

```{r}
data(mtcars)
kmeans_result <- kmeans(mtcars %>% select(mpg, hp, wt), centers = 3)
hc <- hclust(dist(mtcars %>% select(mpg, hp, wt)))

```
